---
title: "Catwalk Analysis"
author: "Tom"
date: "February 22, 2018"
output: html_document
---

```{r, setup, message = F, warning=F}
library(tidyverse)
library(broom)
library(skimr)
library(purrr)
library(agricolae)
library(emmeans)
library(stringr)
library(writexl)
```


```{r}
# have to set to helmert contrasts
options(contrasts = c("contr.helmert", "contr.poly"))

# custom function to generate a tidy type 3 ANOVA
type_3_aov <- function(x, dataframe) {
  broom::tidy(car::Anova(lm(x ~ sex * age * gt, 
                            data=dataframe), 
                         type=3)) %>% 
    mutate(p.value = round(p.value, 3))
}

# custom function to generate a tukey post-hoc
post_hoc_lm <- function(x, dataframe){
  summary(lsmeans::lsmeans(lm(x ~ sex * age * gt, data = dataframe), 
                           pairwise ~ sex:age:gt, 
                           adjust = "tukey"))$contrasts %>% 
    mutate(p.value = round(p.value, 3)) %>% 
    filter(p.value < 0.051)
}

```

```{r, eval = F, echo = F}
# purr functions for calculating all things of interest at once

# this works
catstat_df %>% 
  select(run_speed:run_max_variation) %>% 
  map(~ car::Anova(lm(.x ~ catstat_df$sex * catstat_df$age * catstat_df$gt))) %>% 
  map(~ broom::tidy(.)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5)))
  
# this works!  
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ lm(.x ~ catstat_df$sex * catstat_df$age * catstat_df$gt)) %>% 
  map(~ lsmeans::lsmeans(.x, pairwise ~ catstat_df$sex:catstat_df$age:catstat_df$gt, adjust = "tukey")$contrasts) %>% 
  map(~ tidy(.x)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5))) %>% 
  map(~ filter(.x, p.value < 0.051))

#this works!
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map( ~ pairwise.t.test(.x,
                  catstat_df$sex:catstat_df$age:catstat_df$gt, 
                  p.adj = "none")) %>% 
  map( ~ tidy(.x)) %>% 
  map( ~ mutate(.x, p.value = round(p.value, 5))) %>%
  map( ~ filter(.x, p.value < 0.051))

```




```{r, read in data}
raw_catwalk <- read_csv("catwalk_data.csv")
```


```{r}
skim(raw_catwalk)
head(raw_catwalk)
```

```{r}
cat_df <- raw_catwalk %>% 
  filter(gt != "1" & uid != 928211 & uid != 841350) %>% 
  mutate(age = factor(age, labels = c("4 Month", "10 Month", "17 Month")),
         gt = factor(gt, labels = c("GCLM +/+", "GCLM -/-")))
```
Mental check
```{r}
cat_df %>% 
  filter(age == "17 Month" & run_speed > 70) %>% 
  print()
```

Check distributions of the data.
```{r}
ggplot(cat_df, aes(x = run_speed, fill = age)) + geom_density(alpha = 0.5)
ggplot(cat_df, aes(x = age, y = run_speed, fill = age)) + geom_boxplot() + facet_grid(~sex)
```

Custom function to calculate standard error (SE)
```{r}

se <- function(x) {
  sd(x)/sqrt(length(x))
}
```

Summarize at multiple columns and apply multiple functions, in this case we calculate the mean and se for all of our outcome measures!

```{r}
cat_sum_df <- cat_df %>% 
  group_by(sex, age, gt) %>% 
  summarize_at(vars(run_speed:hind_stepcyc), funs(mean, se)) 
cat_sum_df

```



```{r, eval = F}
cat_df2 <- data.frame(cat_sum_df)
viz_func <- function(data, x, y, z, y.min){
  ggplot(data, 
       aes(x = x, y = y, fill = z, group = z)) + 
  geom_errorbar(aes(ymin = y, 
                    ymax = y + y.min), 
                position = "dodge", width = 0.5) +
  geom_bar(color = "black", stat = "identity", width = 0.7, 
           position = "dodge")
}

viz_func(cat_df2, age, front_bos_mean, gt, run_speed_se)
```



```{r}
head(cat_sum_df)
```







run_speed graph

```{r}
pd_bar <- position_dodge(width = 0.9)
ggplot(cat_sum_df, 
       aes(x = age, y = run_speed_mean, fill = gt, group = gt)) + 
  geom_errorbar(aes(ymin = run_speed_mean, 
                    ymax = run_speed_mean + run_speed_se), position = pd_bar, width = 0.5) +
  geom_bar(color = "black", stat = "identity", width = 0.7, position = pd_bar) +
  facet_grid(~sex)
```

bos_front graph

```{r}
ggplot(cat_sum_df, 
       aes(x = age, y = run_speed_mean, fill = gt, group = gt)) + 
  geom_errorbar(aes(ymin = run_speed_mean, 
                    ymax = run_speed_mean + run_speed_se), position = pd_bar, width = 0.5) +
  geom_bar(color = "black", stat = "identity", width = 0.7, position = pd_bar) +
  facet_grid(~sex)


```

This one is working!
```{r,eval = F}
cat_sum_df2 <- cat_df %>% 
  group_by(age, gt, sex) %>% 
  summarize_at(vars(run_speed:hind_stepcyc), funs(mean, se))
```

```{r}
pd_bar <- position_dodge(width = 0.9)

ggplot(data = cat_sum_df, 
       aes(x = age, y = run_speed_mean, fill = gt,  
           group = gt)) +
  geom_errorbar(aes(ymin = run_speed_mean - run_speed_se, 
                    ymax = run_speed_mean + run_speed_se), width = 0.5, position = pd_bar) +
  geom_bar(color = "black", stat = "identity", width = 0.7, position = pd_bar) +
  
  facet_grid(~sex) +
  theme_bw() +
  scale_fill_manual(values = c("white","black")) +
  theme(legend.position = "NULL",
        legend.title = element_blank(),
        axis.title = element_text(size = 20),
        legend.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12))
```


```{r}
catstat_df <- raw_catwalk %>% 
  filter(gt != "1" & uid != 928211 & uid != 841350) %>% 
  mutate(age = factor(age, labels = c("YG", "AD", "OLD")),
         gt = factor(gt, labels = c("WT", "KO")),
         sex = factor(sex, labels = c("M", "F")))

catstat_df %>% 
  group_by(sex) %>% 
  count()
```
Speed ANOVA
```{r}
speed_aov <- aov(run_speed~sex*age*gt, data = catstat_df)
summary(speed_aov)
TukeyHSD(speed_aov, which = 'sex:age:gt')
# call and save the pair.t.test
speed_lsd <- tidy(pairwise.t.test(catstat_df$run_speed,
                             catstat_df$sex:catstat_df$age:catstat_df$gt, 
                             p.adj = "none")) %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
                
speed_lsd

# type 3 ANOVA and corresponding post-hoc
speed_aov_3 <- type_3_aov(catstat_df$run_speed, catstat_df)
speed_aov_3
speed_posthoc <- post_hoc_lm(catstat_df$run_speed, catstat_df)
```

front_bos ANOVA
```{r}
front_bos_aov <- aov(front_bos~age*gt, data = catstat_df)
summary(front_bos_aov)
TukeyHSD(front_bos_aov)

# call and save the pair.t.test
front_bos_lsd <- tidy(pairwise.t.test(catstat_df$front_bos,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
front_bos_lsd
```

hind_bos ANOVA
```{r}
hind_bos_aov <- aov(hind_bos~age*gt, data = catstat_df)
summary(hind_bos_aov)
TukeyHSD(hind_bos_aov)

# call and save the pair.t.test
hind_bos_lsd <- tidy(pairwise.t.test(catstat_df$hind_bos,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
hind_bos_lsd
```

front_bos ANOVA
```{r}
front_bos_aov <- aov(front_bos~age*gt, data = catstat_df)
summary(front_bos_aov)
TukeyHSD(front_bos_aov)

# call and save the pair.t.test
front_bos_lsd <- tidy(pairwise.t.test(catstat_df$front_bos,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
front_bos_lsd
```

front_strlen ANOVA
```{r}
front_strlen_aov <- aov(front_strlen~age*gt, data = catstat_df)
summary(front_strlen_aov)
TukeyHSD(front_strlen_aov)

# call and save the pair.t.test
front_strlen_lsd <- tidy(pairwise.t.test(catstat_df$front_strlen,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
front_strlen_lsd
```

hind_strlen ANOVA
```{r}
hind_strlen_aov <- aov(hind_strlen~age*gt, data = catstat_df)
summary(hind_strlen_aov)
TukeyHSD(hind_strlen_aov)

# call and save the pair.t.test
hind_strlen_lsd <- tidy(pairwise.t.test(catstat_df$hind_strlen,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
hind_strlen_lsd

```
front_swingspd ANOVA
```{r}
front_swingspd_aov <- aov(front_swingspd~age*gt, data = catstat_df)
summary(front_swingspd_aov)
TukeyHSD(front_swingspd_aov)

# call and save the pair.t.test
front_swingspd_lsd <- tidy(pairwise.t.test(catstat_df$front_swingspd,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
front_swingspd_lsd
```

hind_swingspd ANOVA
```{r}
hind_swingspd_aov <- aov(hind_swingspd~age*gt, data = catstat_df)
summary(hind_swingspd_aov)
TukeyHSD(hind_swingspd_aov)

# call and save the pair.t.test
hind_swingspd_lsd <- tidy(pairwise.t.test(catstat_df$hind_swingspd,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
hind_swingspd_lsd
```

front_stepcyc ANOVA
```{r}
front_stepcyc_aov <- aov(front_stepcyc~age*gt, data = catstat_df)
summary(front_stepcyc_aov)
TukeyHSD(front_stepcyc_aov)

# call and save the pair.t.test
front_stepcyc_lsd <- tidy(pairwise.t.test(catstat_df$front_stepcyc,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
front_stepcyc_lsd
```

hind_stepcyc ANOVA
```{r, eval = F, echo = F}
hind_stepcyc_aov <- aov(hind_stepcyc~age*gt, data = catstat_df)
summary(hind_stepcyc_aov)
TukeyHSD(hind_stepcyc_aov)

# call and save the pair.t.test
hind_stepcyc_lsd <- tidy(pairwise.t.test(catstat_df$hind_stepcyc,
                             catstat_df$age:catstat_df$gt, 
                             p.adj = "none"))  %>% 
              mutate(p.value = round(p.value, digits = 3)) %>% 
              filter(p.value < 0.051)
hind_stepcyc_lsd

```

```{r}
# calculate catwalk ANOVAs and write to Excel
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ car::Anova(lm(.x ~ sex * age * gt, data = catstat_df))) %>% 
  map(~ broom::tidy(.)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5)), .id = 'source') %>% 
  write_xlsx(., "catwalk_anovas.xlsx")

# calculate LSD post-hocs and write to excel
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map( ~ pairwise.t.test(.x,
                  catstat_df$sex:catstat_df$age:catstat_df$gt, 
                  p.adj = "none")) %>% 
  map( ~ tidy(.x)) %>% 
  map( ~ mutate(.x, p.value = round(p.value, 5))) %>%
  map( ~ filter(.x, p.value < 0.051)) %>% 
  write_xlsx(., "catwalk_lsds.xlsx")
```


```{r, echo = F, eval = F}
# purr functions for calculating all things of interest at once

# this works
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ car::Anova(lm(.x ~ catstat_df$sex * catstat_df$age * catstat_df$gt))) %>% 
  map(~ broom::tidy(.)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5)))

catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ car::Anova(lm(.x ~ sex * age * gt, data = catstat_df))) %>% 
  map(~ broom::tidy(.)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5)), .id = 'source') %>% 
  write_xlsx(., "catwalk_anovas.xlsx")

# this works!  
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ lm(.x ~ catstat_df$sex * catstat_df$age * catstat_df$gt)) %>% 
  map(~ lsmeans::lsmeans(.x, pairwise ~ catstat_df$sex:catstat_df$age:catstat_df$gt, adjust = "tukey")$contrasts) %>% 
  map(~ tidy(.x)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5))) %>% 
  map(~ filter(.x, p.value < 0.051))

catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ lm(.x ~ sex * age * gt, data = catstat_df)) %>% 
  map(~ lsmeans::lsmeans(.x, pairwise ~ sex:age:gt, adjust = "tukey")$contrasts) %>% 
  map(~ tidy(.x)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5))) %>% 
  map(~ filter(.x, p.value < 0.051))

catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map(~ lm(.x ~ sex * age * gt, data = catstat_df)) %>% 
  map(~ lsmeans::lsmeans(.x, pairwise ~ sex:age:gt, adjust = "tukey", data = catstat_df)$contrasts) %>% 
  map(~ tidy(.x)) %>% 
  map(~ mutate(.x, p.value = round(p.value, 5))) %>% 
  map(~ filter(.x, p.value < 0.051))

#this works!
catstat_df %>% 
  select(run_speed:hind_stepcyc) %>% 
  map( ~ pairwise.t.test(.x,
                  catstat_df$sex:catstat_df$age:catstat_df$gt, 
                  p.adj = "none")) %>% 
  map( ~ tidy(.x)) %>% 
  map( ~ mutate(.x, p.value = round(p.value, 5))) %>%
  map( ~ filter(.x, p.value < 0.051)) %>% 
  write_xlsx(., "catwalk_lsds.xlsx")

lsd_post$run_speed

```
